---
title: "Exploratory data analysis"
author: "Thierry Onkelinx"
date: "4-3-2021"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ladybird)
library(plotly)
library(INBOtheme)
min_occurrences <- 1000
min_species <- 3
```

## Selection criteria for relevant data

- at least `r min_occurrences` occurrences per species.
- at least `r min_species` observed species per combination of year and location.

We check both criteria twice.
Removing year - location combinations with less than `r min_species` can reduce the number of occurrences of some species.
Removing species with less than `r min_occurrences` occurrences can reduce the number of species for some year - location combinations.

```{r load-data}
relevant <- load_relevant(
  min_occurrences = min_occurrences, min_species = min_species
)
```

```{r locations}
p <- relevant %>%
  count(year) %>%
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  ggtitle("number of relevant locations") +
  theme(axis.title = element_blank())
ggplotly(p, dynamicTicks = TRUE)
```

```{r number-species}
p <- relevant %>%
  pivot_longer(
    c(-year, -location), names_to = "species", values_to = "observed"
  ) %>%
  filter(observed  == 1) %>%
  count(year, location) %>%
  group_by(year) %>%
  summarise(
    mean = mean(n),
    lcl = quantile(n, 0.05),
    ucl = quantile(n, 0.95)
  ) %>%
  ggplot(aes(x = year, y = mean, ymin = lcl, ymax = ucl)) +
  geom_pointrange() +
  ggtitle("Number of species per location (average, 5% and 95% quantile)") + 
  theme(axis.title = element_blank())
ggplotly(p, dynamicTicks = TRUE)
```

```{r high-observed}
relevant %>%
  pivot_longer(
    c(-year, -location), names_to = "species", values_to = "observed"
  ) %>%
  group_by(year, species) %>%
  summarise(proportion = mean(observed), .groups = "drop") -> proportions
proportions %>%
  filter(year == max(year), species != "Hrma") %>%
  filter(proportion >= median(proportion)) -> top_species
p <- proportions %>%
  filter(species %in% c(top_species$species, "Hrma")) %>%
  mutate(Harmonia  = species == "Hrma") %>%
  ggplot(aes(x = year, y = proportion, colour = species, linetype = Harmonia)) +
  geom_line() +
  theme(axis.title = element_blank()) +
  ggtitle("Evolution of species with highest observed proportion in 2020")
ggplotly(p, dynamicTicks = TRUE) %>%
  layout(yaxis = list(tickformat = "%"))
```

```{r low-observed}
p <- proportions %>%
  filter(!species %in% top_species$species) %>%
  mutate(Harmonia  = species == "Hrma") %>%
  ggplot(aes(x = year, y = proportion, colour = species, linetype = Harmonia)) +
  geom_line() +
  theme(axis.title = element_blank()) +
  ggtitle("Evolution of species with lowest observed proportion in 2020")
ggplotly(p, dynamicTicks = TRUE) %>%
  layout(yaxis = list(tickformat = "%"))
```

