---
title: "Results"
author: "Thierry Onkelinx"
date: "8-3-2021"
output: html_document
params:
  models:
    label: "Select a base model"
    value: "../../Analysis/models/base/harm_axyr_1000_3.rds"
    input: file
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
library(ladybird)
library(tidyverse)
library(plotly)
library(scales)
library(INBOtheme)
```

```{r load-base-models}
dirname(params$models) %>%
  list.files(full.names = TRUE) %>%
  map(readRDS) -> base_models
species <- map_chr(base_models, "species")
waic <- map_dbl(base_models, "waic")
map_lgl(base_models, "first_order") %>%
  ifelse(1, 2) -> rw_order
map(base_models, "hyperpar") %>%
  map(rownames_to_column, var = "parameter") %>%
  map2(species, ~mutate(.x, species = .y)) %>%
  map2(rw_order, ~mutate(.x, rw_order = .y)) %>%
  bind_rows() %>%
  mutate(model = "base") -> hyperpar
map(base_models, "trend") %>%
  map2(species, ~mutate(.x, species = .y)) %>%
  map2(rw_order, ~mutate(.x, rw_order = .y)) %>%
  bind_rows() %>%
  mutate(model = "base") -> trend
```

```{r load-secondary-models}
dirname(params$models) %>%
  file.path("..", "secondary") %>%
  list.files(full.names = TRUE) %>%
  map(readRDS) -> secondary_models
sec_species <- map_chr(secondary_models, "species")
sec_waic <- map_dbl(secondary_models, "waic")
map_lgl(secondary_models, "first_order") %>%
  ifelse(1, 2) -> sec_rw_order
map(secondary_models, "fixed") %>%
  map(rownames_to_column, var = "parameter") %>%
  map2(sec_species, ~mutate(.x, species = .y)) %>%
  map2(sec_rw_order, ~mutate(.x, rw_order = .y)) %>%
  bind_rows() -> sec_fixed
map(secondary_models, "hyperpar") %>%
  map(rownames_to_column, var = "parameter") %>%
  map2(sec_species, ~mutate(.x, species = .y)) %>%
  map2(sec_rw_order, ~mutate(.x, rw_order = .y)) %>%
  bind_rows() %>%
  mutate(model = "occurrence") -> sec_hyperpar
map(secondary_models, "trend") %>%
  map2(sec_species, ~mutate(.x, species = .y)) %>%
  map2(sec_rw_order, ~mutate(.x, rw_order = .y)) %>%
  bind_rows() %>%
  mutate(model = "occurrence") -> sec_trend
```

```{r combined}
trend %>%
  mutate(type = "") %>%
  bind_rows(sec_trend) %>%
  mutate(prediction = interaction(model, type, sep = " ")) -> all_trend
hyperpar %>%
  bind_rows(sec_hyperpar) -> all_hyperpar
```

# Compare models

- **base**: intercept + random walk along year + spatiotemporal field
- **occurrence**: base model + logit predicted probability of _H. axyridis_ as fixed effect

- model **1**: first order random walk
- model **2**: second order random walk

```{r waic}
data.frame(species, waic, rw_order = rw_order, model = "base") %>%
  bind_rows(
    data.frame(
      species = sec_species, waic = sec_waic, rw_order = sec_rw_order,
      model = "occurrence"
    )
  ) %>%
  group_by(species) %>%
  mutate(
    waic = waic - min(waic),
    model = paste(model, rw_order)
  ) %>%
  select(-rw_order) %>%
  pivot_wider(names_from = model, values_from = waic) %>%
  kable(
    digits = 1,
    caption = 
  "Difference in WAIC compared to the model with lowest WAIC for each species."
  )
```


# Interactions with _H. axyridis_

```{r sec-fixed, fig.cap = "Parameter estimate when using the (logit) predicted occurrence of H. axyridis as fixed effect covariate."}
sec_fixed %>%
  filter(parameter == "secondary") %>%
  mutate(
    species = reorder(species, mean),
    order = factor(rw_order)
  ) %>%
  ggplot(
    aes(
      x = mean, xmin = `0.025quant`, xmax = `0.975quant`, y = species,
      colour = order
    )
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_errorbarh(position = position_dodge(width = 0.5)) +
  geom_point(position = position_dodge(width = 0.5)) +
  theme(axis.title = element_blank())
```

## General trends

- **occurrence median**: prediction using the median predicted occurrence of _H. axyridis_ for that year
- **occurrence without**: prediction assuming the predicted occurrence of _H. axyridis_ is 0.1%

```{r trends-adal-bipu}
plot_trend <- function(this_species) {
  all_trend %>%
    filter(species  == this_species, type %in% c("median", "without", "")) %>%
    ggplot(aes(x = year, y = mean, ymin = lcl, ymax = ucl)) +
    geom_ribbon(aes(fill = prediction, colour = prediction), alpha = 0.3, linetype = 3) +
    geom_line(aes(colour = prediction)) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(axis.title = element_blank()) +
    ggtitle(this_species) +
    facet_wrap(~rw_order)
}
plot_trend("Adal_bipu")
```
```{r trends-adal-dece}
plot_trend("Adal_dece")
```

```{r trends-calv-dece}
plot_trend("Calv_dece")
```

```{r trends-calv-quat}
plot_trend("Calv_quat")
```

```{r trends-cocc-sept}
plot_trend("Cocc_sept")
```


```{r trends-cocc-quin}
plot_trend("Cocc_quin")
```

```{r trends-exoc-quad}
plot_trend("Exoc_quad")
```

```{r trends-haly-sede}
plot_trend("Haly_sede")
```

```{r trends-harm-axyr}
plot_trend("Harm_axyr")
```

```{r trends-harm-quad}
plot_trend("Harm_quad")
```

```{r trends-hipp_vari}
plot_trend("Hipp_vari")
```

```{r trends-oeno-cong}
plot_trend("Oeno_cong")
```

```{r trends-prop-quat}
plot_trend("Prop_quat")
```

```{r trends-psyl-vigi}
plot_trend("Psyl_vigi")
```

```{r trends-tytt-sede}
plot_trend("Tytt_sede")
```

## Spatio-temporal fields

```{r sec-range}
all_hyperpar %>%
  filter(parameter == "Range for site") %>%
  mutate(
    species = reorder(species, mean),
    order = factor(rw_order)
  ) %>%
  ggplot(
    aes(
      x = mean, xmin = `0.025quant`, xmax = `0.975quant`, y = species,
      colour = model, linetype = order, shape = order
    )
  ) +
  geom_errorbarh(position = position_dodge(width = 0.5)) +
  geom_point(position = position_dodge(width = 0.5)) +
  scale_x_continuous(limits = c(0, NA)) +
  ggtitle("Range of the spatio-temporal field in km") +
  theme(axis.title = element_blank())
```

```{r sec-stdev}
all_hyperpar %>%
  filter(parameter == "Stdev for site") %>%
  mutate(
    species = reorder(species, mean),
    order = factor(rw_order)
  ) %>%
  ggplot(
    aes(
      x = mean, xmin = `0.025quant`, xmax = `0.975quant`, y = species,
      colour = model, linetype = order, shape = order
    )
  ) +
  geom_errorbarh(position = position_dodge(width = 0.5)) +
  geom_point(position = position_dodge(width = 0.5)) +
  scale_x_continuous(limits = c(0, NA)) +
  ggtitle("Stdev of the spatio-temporal field in km") +
  theme(axis.title = element_blank())
```

```{r sec-rho}
all_hyperpar %>%
  filter(parameter == "GroupRho for site") %>%
  mutate(
    species = reorder(species, mean),
    order = factor(rw_order)
  ) %>%
  ggplot(
    aes(
      x = mean, xmin = `0.025quant`, xmax = `0.975quant`, y = species,
      colour = model, linetype = order, shape = order
    )
  ) +
  geom_errorbarh(position = position_dodge(width = 0.5)) +
  geom_point(position = position_dodge(width = 0.5)) +
  ggtitle("Rho of the spatio-temporal field in km") +
  theme(axis.title = element_blank())
```
